\sysfull

### 2018-08-05 paper/asplos19/implementation.tex
revise implemenation

We implemented \sys's programming and execution model as a runtime library and API that a programmer can use to make a plain C program intermittence-safe.

\subsection{Application Programming Interface}
\label{sec:coala_api}

Coala's API adds only eight syntactic constructs to a C-based language, summarized
in Table~\ref{table:coala_api}.

%\texttt{TASK} is a function annotation that declares a function to be a
%task. \texttt{NEXT\_TASK}, indicates the task to run next, after an
%executing task completes.  \texttt{PV} declare a new protected variable.
%The programmer uses \texttt{RP} and \texttt{WP}, to read and write protected
%respectively.  \texttt{SM} identifies protected \texttt{struct} fields
%and tells \sys to ensure that the struct fields are page-aligned in memory.
%The programmer is required to explicitly call two API functions at the start of
%{\tt main()}: \texttt{INIT}, which does global initialization upon first
%boot and \texttt{RUN} which hands the control to \sys's task manager to
%resume task execution.

\begin{table}
\centering
\begin{tabular}{| r | p{0.67\columnwidth} |}
	\hline
	{Method} & {Arguments} \\
	\hline\hline
	\texttt{INIT}($t$) & $t \in \mathbf{T}$: scheduled task on first boot \\
	\hline
	\texttt{RUN}() & --- \\
	\hline
	\texttt{TASK}($t$, $w_t$) & $t \in \mathbf{T}$: task name, $w_t$: weight of task $t$ \\
	\hline
	\texttt{NEXT\_TASK}($t$) & $t \in \mathbf{T}$ : task to be run next \\
	\hline
	\texttt{PV}($p$, $v$ [, $s$]) & $p$: variable type, $v \in \mathbf{V}$: new protected variable name, $s$: array size \\
	\hline
	$u$ := \texttt{RP}($v$) & $v \in \mathbf{V}$: protected variable to be read, $u$: destination operand \\
	\hline	
	\texttt{WP}($v$) := $u$ &  $v \in \mathbf{V}$: protected variable to be written, $u$: source operand \\
	\hline
	\texttt{SM}($p$, $m$ [, $s$]) & $p$: \texttt{struct} member's type, $m$: \texttt{struct} member's name, $s$: array size \\
	\hline
\end{tabular}
\caption{API summary; $\mathbf{T}$: set of all tasks, $\mathbf{V}$: set of all protected variables, $[, s]$: optional argument.}
\label{table:coala_api}
\end{table}

%\subsection{API Details}

\noindent\textbf{New Tasks.} The \texttt{TASK} annotation on a function
declaration statically allocates a non-volatile constant variable holding a
task's weight and declares that the function is a task.\\
%
\noindent\textbf{Task Transitions.} A call to \texttt{NEXT\_TASK} marks the task to be executed at the 
end of the current one task and it can be invoked along
any control path to dynamically determine the next task.\\
%
\noindent\textbf{Protected Variables.} The \texttt{PV} annotation on a
variable statically allocates a protected non-volatile variable. The
variable must then be accessed with the \texttt{RP} and \texttt{WP} API methods at
runtime to ensure correct operation.\\
%
\noindent\textbf{Initialization.} The behavior of the API method
\texttt{INIT} is very similar to \texttt{NEXT\_TASK}, with the
addition of performing preliminary kernel initializations, including hardware setup.\\
%
\noindent\textbf{Execution.} The programmer passes control to \sys's task scheduler 
by calling \texttt{RUN} after device initialization. 
%
First the scheduler updates the coalescing target using the coalescing
strategy.  Second, the scheduler resumes a commit, if one is in progress. Third,
the scheduler clears the list of dirty shadow pages. The scheduler then 
sets the program counter to the next task to run, which \sys tracks in  
non-volatile memory.  
%
Before executing the task, \sys checks whether there is a partially committed
task to resume, which requires \sys to restore volatile state, including the program
counter.  
%
If there is no in-progress, partially committed task, \sys starts executing and
coalescing tasks.

@@

Page tagging imposes a data alignment requirement.
Page size $S$ must be a power of two. 
%
Pages must be aligned to an $S$-byte boundary for efficient memory access. 
%
We experimented with 32-, 64-, 128- and 256\,B pages, an 8\,KB non-volatile \texttt{\underline{shadow}} buffer and
an 8\,KB non-volatile \texttt{\underline{private}}, and a \texttt{working} buffer
of 1\,KB.
%
\subsection{Coalescing and Commit}
%
\sys uses a two-phase commit process to atomically commit pages despite
possible power interruptions during commit.  When a coalesced task completes,
\sys backs up dirty pages in volatile working memory to the non-volatile
shadow buffer using DMA. 
%
If power fails while backing up these pages, commit fails and the execution
resumes from the last committed point.  After backing up pages from working
memory to the shadow buffer, \sys sets a flag that indicates an on-going atomic
commit.  At this point, commit is irrevocable and will continue until it
completes despite power failures.   
%
Each page is double buffered, with one copy always in the shadow page buffer
and one copy always in the main memory store.  \sys uses double buffering to
commit dirty pages efficiently by swapping which buffer is which, without
copying any data.
%
\subsection{Task Downscaling}
%
\sys's task downscaling preserves forward progress by partially committing an
ongoing, non-terminating task. 
%
Partial commit relies on the existence of a hardware timer, which is a very
common feature in a typical MCU and consumes very little energy (i.e.,
$\approx$3\,$\mu$A/MHz~\cite{msp430datasheet}).

>>>>

We implemented \sys's programming and execution model as a runtime library and API that a programmer can use to make a plain C program intermittence-safe.
%
\subsection{Application Programming Interface}
\label{sec:coala_api}
%
Coala's API adds only eight syntactic constructs to a C-based language, summarized
in Table~\ref{table:coala_api}.
%
\begin{table}
\centering
\begin{tabular}{| r | p{0.67\columnwidth} |}
	\hline
	{Method} & {Arguments} \\
	\hline\hline
	\texttt{INIT}($t$) & $t \in \mathbf{T}$: scheduled task on first boot \\
	\hline
	\texttt{RUN}() & --- \\
	\hline
	\texttt{TASK}($t$, $w_t$) & $t \in \mathbf{T}$: task name, $w_t$: weight of task $t$ \\
	\hline
	\texttt{NEXT\_TASK}($t$) & $t \in \mathbf{T}$ : task to be run next \\
	\hline
	\texttt{PV}($p$, $v$ [, $s$]) & $p$: variable type, $v \in \mathbf{V}$: new protected variable name, $s$: array size \\
	\hline
	$u$ := \texttt{RP}($v$) & $v \in \mathbf{V}$: protected variable to be read, $u$: destination operand \\
	\hline	
	\texttt{WP}($v$) := $u$ &  $v \in \mathbf{V}$: protected variable to be written, $u$: source operand \\
	\hline
	\texttt{SM}($p$, $m$ [, $s$]) & $p$: \texttt{struct} member's type, $m$: \texttt{struct} member's name, $s$: array size \\
	\hline
\end{tabular}
\caption{API summary; $\mathbf{T}$: set of all tasks, $\mathbf{V}$: set of all protected variables, $[, s]$: optional argument.}
\label{table:coala_api}
\end{table}
%
\noindent\textbf{New Tasks.} The \texttt{TASK} annotation on a function
declaration statically allocates a non-volatile constant variable holding a
task's weight and declares that the function is a task.\\
%
\noindent\textbf{Task Transitions.} A call to \texttt{NEXT\_TASK} marks the task to be executed at the 
end of the current one task and it can be invoked along
any control path to dynamically determine the next task.\\
%
\noindent\textbf{Protected Variables.} The \texttt{PV} annotation on a
variable statically allocates a protected non-volatile variable. The
variable must then be accessed with the \texttt{RP} and \texttt{WP} API methods at
runtime to ensure correct operation.\\
%
\noindent\textbf{Initialization.} The behavior of the API method
\texttt{INIT} is very similar to \texttt{NEXT\_TASK}, with the
addition of performing preliminary kernel initializations, including hardware setup.\\
%
\noindent\textbf{Execution.} The programmer passes control to \sys's task scheduler 
by calling \texttt{RUN} after device initialization. 
%
First the scheduler updates the coalescing target using the coalescing
strategy.  Second, the scheduler resumes a commit, if one is in progress. Third,
the scheduler clears the list of dirty shadow pages. The scheduler then 
sets the program counter to the next task to run, which \sys tracks in  
non-volatile memory.  
%
Before executing the task, \sys checks whether there is a partially committed
task to resume, which requires \sys to restore volatile state, including the program
counter.  
%
If there is no in-progress, partially committed task, \sys starts executing and
coalescing tasks.

@@

Page tagging imposes a data alignment requirement.
Page size $S$ must be a power of two. 
%
Pages must be aligned to an $S$-byte boundary for efficient memory access. 
%
We experimented with 32-, 64-, 128- and 256\,B pages, an 8\,KB non-volatile \texttt{\underline{shadow}} buffer and
an 8\,KB non-volatile \texttt{\underline{private}}, and a \texttt{working} buffer
of 1\,KB.
%
\subsection{Coalescing and Commit}
%
The coalescing algorithm starts with a random guess about the size of the coalescing task, in our implementation the target is initialized to four. It then refines its guess based on the number of the static-tasks being successfully executed.
%
When a coalesced task completes, \sys save all the modifications in its \texttt{\underline{private}} using DMA. In order to ensure the atomicity of the commit operation all the modified pages are copied to the \texttt{\underline{shadow}} buffer first. The \sys set a flag to indicates an on-going \emph{atomic commit} to update the private pages. For efficiency reasons \sys uses an indirection table to point to the private consistent pages and eliminate the need for moving pages for the second time.
%
\subsection{Task Downscaling}
%
\sys's task downscaling preserves forward progress by partially committing an
ongoing, non-terminating task. 
%
Partial commit relies on the existence of a hardware timer, which is a very
common feature in a typical MCU and consumes very little energy (i.e.,
$\approx$3\,$\mu$A/MHz~\cite{msp430datasheet}).
